WSClean: an implementation of a fast, generic wide-field imager

### 2014-02-18 articles/2014/Article-WSClean/wsclean.tex
Slight updates

TODO: add big-ass full-sky image

@@

Instead of imaging the whole sky, the image can be recentred from $(l,m)$ to $(\hat{l},\hat{m})=(l+\Delta l,m+\Delta m)$ by performing the substitution $(l,m)\leftarrow(\hat{l},\hat{m})$ in Eq.~\eqref{eq:wstacking}:
%When the Fourier transform is performed, the $w$ term has been corrected (either by $w$-stacking or by $w$-projection). Therefore, if $U(u, v)$ is the input to the 2-dimensional Fourier transform, the transform is described by
\begin{align}\notag
\frac{I'(\hat{l}, \hat{m})\left(w_{\max} - w_{\min}\right)}{\sqrt{1-\hat{l}^2-\hat{m}^2}} = \int\limits_{w_{\min}}^{w_{\max}} e^{2\pi i w(\sqrt{1-\hat{l}^2-\hat{m}^2}-1)} \times \\
\iint e^{2 \pi i \left(u\Delta l + v\Delta m\right)} V(u,v,w) e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
In words, recentering an image involves accounting for the position shift during the $w$-correction and final scaling, and shifting the visibilities in phase by multiplication with $e^{2 \pi i \left(u\Delta l + v\Delta m\right)}$ prior to the imaging. This essentially is a method of implementing the $w$-snapshots algorithm as described by \citet{widefield-imaging-ska-cornwell}. By doing the inverse corrections during prediction, such a shifted visibility set can be cleaned with Cotton-Schwab iterations similar to a non-shifted set.

A recentred image is in the projection of the zenith tangent plane, and will need an additional regridding step before multiple snapshots can be added together. A recentred image can be stored in a Flexible Image Transport System (FITS) file with the slant orthographic (SIN) projection normally used in interferometric imaging \citep{fits-coordinates-2002}, by setting the centre of tangent projection with the \texttt{CRPIXi} keyword \citep{wcs-in-fits}. Common viewers such as Kvis and DS9 support such FITS files and display their coordinates correctly. An example of the difference in projection is given in Fig.~\ref{fig:vela-projection-example}, which displays a 13~min MWA observation imaged with WSClean. Cleaning an image that is in a different projection yields slightly different results, but qualitativily the two images are clearly of equal accuracy. The wall-clock time for imaging is 60 min and 41 min for respectively normal projection and the recentred image. Of the 41 min, 3 min is spent on phase rotating the visibilities.

@@

TODO: I just realized that the original definition of $w$-snapshots might be Natasha's way of fixing $w$-values in MWACS. In that case, these methods are new and it might be good to distinguish this.

@@

For the WSClean time-cost function $t_\textrm{WSClean}$ we find 
\begin{align}\label{eq:wsclean-computing-cost}
& t_\textrm{WSClean} (w_{\max},N_\textrm{Mvis},N_\textrm{freq},N_\textrm{kpix}) = \\ \notag%
& N_\textrm{freq} \left( 0.452 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} (w_{\max} + 0.611) + 244 N_\textrm{Mvis} \right)
\end{align}
and for CASA we find
\begin{align}\label{eq:casa-computing-cost}
& t_\textrm{CASA}(w_{\max},N_\textrm{vis},N_\textrm{freq},N_\textrm{pix}) = \\ \notag%
& N_\textrm{freq} \left(1.03 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} + 5.93 N_\textrm{Mvis} (w_{\max} + 1.71)^2\right) + 172.
\label{eq:casa-computing-cost}%
\end{align}
Parameter $w_{\max}$ is the maximum $w$ value and is estimated with
\begin{equation}
 w_{\max} = \frac{1}{\lambda}\left[D \sin \textrm{ZA} + \max \Delta z \cos \textrm{ZA} \left(1.0 - \cos(\frac{1}{2}\textrm{FOV})\right)\right],
\end{equation}
where $D$ is the maximum baseline length, $\max \Delta z$ is the maximum height difference between antennas and $\lambda$ is the wavelength, all in meters. These functions follow the trend of the measurements well, with an absolute error of 14.7\% and 20.7\% for respectively the WSClean and CASA functions. (TODO $w_{\max}$ is slightly different for CASA)

>>>>

TODO: add big full-sky image

@@

Instead of imaging the whole sky, the image can be recentred from $(l,m)$ to $(\hat{l},\hat{m})=(l+\Delta l,m+\Delta m)$ by performing the substitution $(l,m)\leftarrow(\hat{l},\hat{m})$ in Eq.~\eqref{eq:wstacking}:
\begin{align}\notag
\frac{I'(\hat{l}, \hat{m})\left(w_{\max} - w_{\min}\right)}{\sqrt{1-\hat{l}^2-\hat{m}^2}} = \int\limits_{w_{\min}}^{w_{\max}} e^{2\pi i w(\sqrt{1-\hat{l}^2-\hat{m}^2}-1)} \times \\
\iint e^{2 \pi i \left(u\Delta l + v\Delta m\right)} V(u,v,w) e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
In words, recentering an image involves accounting for the position shift during the $w$-correction and final scaling, and shifting the visibilities in phase by multiplication with $e^{2 \pi i \left(u\Delta l + v\Delta m\right)}$ prior to the imaging. This essentially is a method of implementing the $w$-snapshots algorithm as described by \citet{widefield-imaging-ska-cornwell}. However, instead of creating warped images by performing the FFT over a tilted plane in $uvw$-space, our method transforms planes with constant $w$-values. The benefit of our method is that the resulting image has a circular synthesized beam, i.e., the resolution in $l$ and $m$ directions matches the intrinsic resolution of the instrument, which is desirable for cleaning. Our method is also relatively easy to implement. It has a similar computational cost compared to the $w$-snapshot algorithm, although the required number of $w$-layers has a different dependency on zenith angle, field of view and the non-coplanarity.

By doing the inverse corrections during prediction, a recentred visibility set can be cleaned with Cotton-Schwab iterations similar to a non-recentred set. The PSF used during minor cleaning iterations is created by multiplying the weights with the recentering corrections, such that the PSF represents a source in the middle of the image.

A recentred image is in the projection of the zenith tangent plane, and will need an additional regridding step before multiple snapshots can be added together. A recentred image can be stored in a Flexible Image Transport System (FITS) file with the slant orthographic (SIN) projection normally used in interferometric imaging \citep{fits-coordinates-2002}, by setting the centre of tangent projection with the \texttt{CRPIXi} keyword \citep{wcs-in-fits}. Common viewers such as Kvis and DS9 support such FITS files and display their coordinates correctly. Unlike warped snapshots, recentred images do not need the generalized-SIN-projection keywords \texttt{PV2\_1} and \texttt{PV2\_2}. An example of the difference in projection between non-recentred and recentred imaging is given in Fig.~\ref{fig:vela-projection-example}, which displays a 13~min MWA observation imaged with WSClean. Cleaning an image that is in a different projection yields slightly different results, but qualitativily the two images are clearly of equal accuracy. The wall-clock time for imaging is 60 min and 41 min for respectively normal projection and the recentred image. Of the 41 min, 3 min is spent on phase rotating the visibilities.

@@

@@

For the WSClean time-cost function $t_\textrm{WSClean}$ we find 
\begin{align}\label{eq:wsclean-computing-cost}
& t_\textrm{WSClean} (w_{\max},N_\textrm{Mvis},N_\textrm{freq},N_\textrm{kpix}) = \\ \notag%
& N_\textrm{freq} \left( 0.452 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} (w_{\max} + 0.611) + 244 N_\textrm{Mvis} \right)
\end{align}
and for CASA we find
\begin{align}\label{eq:casa-computing-cost}
& t_\textrm{CASA}(w_{\max},N_\textrm{vis},N_\textrm{freq},N_\textrm{pix}) = \\ \notag%
& N_\textrm{freq} \left(1.03 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} + 5.93 N_\textrm{Mvis} (w_{\max} + 1.71)^2\right) + 172.
\label{eq:casa-computing-cost}%
\end{align}
Parameter $w_{\max}$ is the maximum $w$-value and is estimated with
\begin{equation}\resizebox{0.45\textwidth}{!}{$
 w_{\max} = \frac{1}{\lambda}\left[(D \sin \textrm{ZA} + \Delta z_{\max} \cos \textrm{ZA} + \xi) \left(1.0 - \cos(\frac{1}{2}\textrm{FOV})\right)\right]$},
\end{equation}%
where $D$ is the maximum baseline length, $\Delta z_{\max}$ is the maximum height difference between antennas and $\lambda$ is the wavelength, all in meters, and $\xi$ is an extra parameter for fitting the CASA measurements, $\xi_\textrm{CASA}=28.4$ and $\xi_\textrm{WSClean}=0$. These functions follow the trend of the measurements well, with an absolute error of 14.7\% and 20.7\% for respectively the WSClean and CASA functions.
